Supplementary Information

### 2020-12-30 doc/manuscript/MEE_manuscript.Rmd
BCI-->BCNM, site table updates
#100

all_GLS = "S9-S18"
GLS_BCI = "S9"
GLS_HKK = "S10"
GLS_SCBI = "S11"
GLS_LDW = "S12"
GLS_HF = "S13"
GLS_ZOF = "S14"
GLS_NE = "S15"
GLS_LT = "S16"
GLS_CB = "S17"
GLS_SC = "S18"

@@

climwin_sameV_sameW_legend = "Figure S5. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S6. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S7. (TMP/TMN at BCI)"

GLS_BCI_legend = "Figure S9. Best GLS models for Barro Colorado Island (Panama)"
GLS_HKK_legend = "Figure S10. Best GLS models for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S11. Best GLS models for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S12. Best GLS models for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S13. Best GLS models for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S14. Best GLS models for Zofin Forest (Czech Republic)"
GLS_NE_legend = "Figure S15. Best GLS models for Niobrara/ Halsey (Nebraska, USA)"
GLS_LT_legend = "Figure S16. Best GLS models for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S17. Best GLS models for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S18. Best GLS models for Scotty Creek (Northwest Territory, Canada)"
```

@@

**Table 1 | Sites included in this analysis**. Here and throughout, sites are ordered by descending mean annual temperature. Additional site information is provided in Table `r site_details`, and species and their sample sizes are detailed in Tables `r species_table`-`r sampling_details_table`. 
![ ](tables_figures/sites_table.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @vlam_temperature_2014; @maxwell_declining_2016; @sniderhan_growth_2016; @alfaro-sanchez_growth_2017; @helcoski_growing_2019 @gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020-1; @aus_de_ar_tree_2018; @bumann_assessing_2019; @samonil_individual-based_2013; @kaspar_species-specific_nodate; @alexander_potential_2019; @finzi_carbon_2020
---

@@

```{r , eval=FALSE, echo=FALSE}
## Tallies for the following paragraph (Dec.21)
#sameV_sameW: LDWp, NMt, NIOt, NIOp, SCBIp, SCt, ZOFp
#sameV_diffW: BCIp, BCIt, CBp, CBt, HFt, HKKp, LDWt, NMp, SCBIt, SCp, ZOFt
#diffV_diffW: HFp, HKKt
```
The three metrics of growth ($RW$, $BAI$, and $\Delta AGB$)  exhibited similar strength of correlation and direction of response to climate variables across the range of potential time windows, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).
In 7 of 20 cases, both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_sameV_sameW`). 
In 11 cases, *climwin* identified the same climate variable but different (often overlapping) time windows (e.g., Fig. `r climwin_sameV_diffW`). 
Finally, in two cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at HKK, precipitation variable group at HF; Figs. `r clim_sens`, `r GLS_HKK`, `r GLS_HF`), *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_diffV_diffW`). 
Henceforth, unless otherwise noted, we focus on the climate sensitivities identified with $RW$ as the growth metric and for the full set of cores (*i.e.*, including those for which $DBH$ could not be reconstructed).

```{r , eval=FALSE, echo=FALSE}
## Tallies for the paragraph below (as of Dec. 22)
#BCI: 1 species with both, 2 with just P
#HKK: 3 species with both, 1 with just T
#SCBI: 11 species with both, 3 with just P (FRNI, FAGR, PIST)
#LDW:  4 species with both,  2  with just P
#HF:  4 species with both
#ZOF:  3 species with both
#NIO:  1 species with both
#LT:  2 species with both
#CB:  7 species with both,  1 with just P (POTR)
#SC:  1 species with both
```
Within the context of the GLS model, which tested for first- and negative second- order linear effects of both a precipitation and a temperature variable, both a precipitation and a temperature variable were included in the top model for 80% of site-species combinations (n=37 of 46; Fig. `r clim_sens`). 
There were eight site-species combinations for which only a precipitation term was significant (2 at BCI, 3 at SCBI, 2 at LDW, and 1 at CB), one for which only a temperature term was significant (CHTA at HKK), and none with no significant climate responses. 
Precipitation and temperature variables were rarely influential over the same time window (exception: LDW).
Below, we summarize the precipitation and temperature variables included in these models and their direction of response.

@@

###Windows selected: 
# BCI: precip dry season (12mo), T dry season (2 mo) 
#HKK: PDF dry season + edges (9mo), T wet season (8 mo)
#SCBI: precip p June - cAug (15 mo), PET current gs (3 mo)
#LDW: precip c June (1 mo), PET current June (1mo)
#HF: PDF pMay (1mo), PET cMar (1mo)
#ZOF: pre pJune-pJul (2mo), TMX Jan-Mar (3mo)
#NIO: wet cMay (1mo), TMP pJune (1mo)
#LT: P pNov-cJul (9mo), TMX cMay-Jul (3mo)
#CB: P cApr-cJul(4mo), TMXpJul-cJul(13mo)
#SC: P pJun-pAug(3mo), TMXcJul (1mo)
```
Responses to precipitation amount ($PPT$) and frequency ($PDF$) were included in the best model for all but one species, and were predominantly positive (Fig. `r clim_sens`).
$PPT$ and $PDF$ were each identified as the top variable in four of the eight sites with both variables available, while $PPT$ was the only option at two (LT and CB).
Optimal time windows most commonly coincided at least partially with the months of most active growth in the current year (Fig. `r clim_sens`; Table `r site_details`):  wet seasons in the tropics (BCI and HKK) or late spring/ early summer in the extratropics (n= 5 of 8).
In the tropics, the long time windows over which precipitation was influential (12 mo at BCI, 9 mo at HKK) also included the majority (BCI) or all (HKK) of the dry season.
At three extratropical sites (LDW, NIO, and CB), the most influential windows were limited to the current spring and early summer, whereas optimal windows at two sites extended back to the previous fall (LT) or summer (SCBI).
Finally, at three sites (HF, ZOF, and SC), precipitation of the previous growing season was the most influential variable. 
Responses to $PPT$ or $PDF$ were predominantly positive, with positive first-order linear terms for all but two species (*Tsuga canadensis* at HF and *Betula papyrifera* at NIO; Fig. `r clim_sens`).
Negative second-order terms were commonly included in the best model (*# of #* with positive first-order terms), generally resulting in a deceleration or decline at the highest levels of precipitation, but occasionally producing a unimodal (e.g., several species at SCBI) or predominantly negative response (e.g., BEAL at HF; Fig. `r clim_sens`). 
At the riparian site (NIO), for which streamflow was included as a candidate variable in the precipitation group, May $PDF$ was selected as the strongest driver but had a modest negative effect on growth, indicating that water availability was not a significant limitation for *Betula papyrifera* at this site.

A temperature variable was included in the best model for all but eight site-species combinations, with predominantly negative responses (*#*% of species; Fig. `r clim_sens`).
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, which were identified by *climwin* as the top temperature-related driver at five and three of the ten sites, respectively, noting that $PET$ was not available for two sites. 
$T_{mean}$ and $T_{min}$ were each identified as the top driver at only one site (NIO and BCI, respectively), noting that the effects of $T_{min}$ at BCI were only marginally significant (Fig. `r clim_sens`).
Optimal time windows most commonly occurred during the peak growing season of the current year (n= 5 of 10 sites), but there were cases where optimal windows occurred during the preceding dry season (BCI), late winter/early spring (HF, ZOF), the previous growing season (NIO), or a 13-month window extending from past to current growing season (CB).
Within the tropics, there was minimal effect of temperature at BCI and a negative effect of wet season $T_{max}$ for three of four species at HKK.
For temperate sites with optimal time windows covering the current and/or past growing season, responses were universally negative (i.e., negative first-order linear or unimodal, peaking at temperatures lower than the long-term mean). 
In contrast, there were positive effects of Jan-March $T_{max}$ for all three species at ZOF and of March $PET$ for *Tsuga Canadensis* at HF, the latter contrasting with a negative response of the other three species at HF (Fig. `r clim_sens`).
At the highest-latitude site (SC), which has undergone rapid warming and permafrost melt, *Picea mariana* responded positively (but with wide 95% CI on the slope) to temperature over the full analysis period (1903-2013); however, responses were predominately positive prior to 1970 and predominantly negative afterwards (Fig. **CREATE AN SI FIGURE!**).

@@

**All three growth metrics, $RW$, $BAI$, and $\Delta AGB$, varied with $DBH$ for most species at all sites (Fig. `r size`).** 
$RW$ varied significantly with $DBH$ for `r round(n_spst_DBH_sig/ n_spst_DBH_analyzed*100)`% of species-site combinations (n= `r n_spst_DBH_sig` of `r n_spst_DBH_analyzed`), and the majority of best models included a significant second-order linear $DBH$ term (n= `r n_spst_DBH_2nd_order`, `r n_spst_DBH_2nd_order_neg` of which were negative; Table **S#**). 
There was substantial variation in these trends, with patterns mixed across both forests and species within a single stand (Fig. `r size`).
On one end of the spectrum, some species exhibited maximum $RW$ at low $DBH$, followed by fairly rapid declines in $RW$ with increasing $DBH$. 
Species following this pattern either had low to at most intermediate shade tolerance (e.g., *Melia azedarach* at HKK, *Juglans nigra* at SCBI, *Acer rubrum* at HF, *Betula papyrifera* at NIO; *Populous tremuloides* at CB; Table `r species_table`) and/or grew in relatively open stands (e.g., both species at LT, *Picea mariana* at SC; Fig. `r size`).
At the other end of the spectrum, some species had low $RW$ at small $DBH$, increased to peak $RW$ at intermediate $DBH$, and subsequently declined.
These included, but were not limited to, shade-tolerant species at sites where they presumably established under closed-canopy conditions (e.g., *Trichilia tuberculata* and *Tetragastris panamensis* at BCI; *Fagus* spp. at SCBI and Žofín, *Acer saccharum* at LDW, *Picea* spp. at Žofín and CB; Table `r species_table`).

@@

When a precipitation variable, a temperature variable, and $DBH$ were all included as candidate variables in the GLS models, typically all three were included in the top model, regardless of the growth metric used.
Climate responses were generally similar to those described above for models without a $DBH$ term, although some of the weaker climate responses were not consistently included in top models (e.g., $T_{min}$ responses at BCI; Figs. `r clim_sens`, `r all_GLS`).
Responses to $DBH$ were as described above (Fig. `r size`).

In general, $DBH$ explained more variation in growth rates than did climate, but its relative importance varied across growth metrics and sites (Figs. `r all_GLS`).
The relative importance of $DBH$ tended to be least for $RW$, intermediate for $BAI$, and highest for $\Delta AGB$ (e.g., at SCBI; Fig. `r GLS_SCBI`).
However, there were exceptions, particularly when $RW$ decreased steeply with $DBH$ (e.g., LT; Fig. `r GLS_LT`).
The relative importance of climate was modest (but still significant) at sites including SCBI (Fig. `r GLS_SCBI`), HF (Fig. `r GLS_HF`), and SC (Fig. `r GLS_SC`), and stronger at sites including BCI (Fig. `r GLS_BCI`) and LT (Fig. `r GLS_LT`).

@@

Main bias would be stand dynamics. This is probably the case at secondary forests (HF, SCBI...). Even within mature forests, light-demanding species that establish in gaps (JACO at BCI, MEAZ at HKK) would tend to experience an increasingly competitive environment.

>>>>

all_GLS = "S9-S18"
GLS_BCNM = "S9"
GLS_HKK = "S10"
GLS_SCBI = "S11"
GLS_LDW = "S12"
GLS_HF = "S13"
GLS_ZOF = "S14"
GLS_NE = "S15"
GLS_LT = "S16"
GLS_CB = "S17"
GLS_SC = "S18"

@@

climwin_sameV_sameW_legend = "Figure S5. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S6. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S7. (TMP/TMN at BCNM)"

GLS_BCNM_legend = "Figure S9. Best GLS models for Barro Colorado Nature Monument (Panama)"
GLS_HKK_legend = "Figure S10. Best GLS models for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S11. Best GLS models for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S12. Best GLS models for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S13. Best GLS models for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S14. Best GLS models for Zofin Forest (Czech Republic)"
GLS_NE_legend = "Figure S15. Best GLS models for Niobrara/ Halsey (Nebraska, USA)"
GLS_LT_legend = "Figure S16. Best GLS models for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S17. Best GLS models for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S18. Best GLS models for Scotty Creek (Northwest Territory, Canada)"
```

@@

**Table 1 | Sites included in this analysis**. Here and throughout, sites are ordered by descending mean annual temperature. Additional site information is provided in Table `r site_details`, and species and their sample sizes are detailed in Tables `r species_table`-`r sampling_details_table`. 
![ ](tables_figures/sites_table.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @vlam_temperature_2014; @maxwell_declining_2016; @sniderhan_growth_2016; @alfaro-sanchez_growth_2017; @helcoski_growing_2019 @gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020-1;  @bumann_assessing_2019; @samonil_individual-based_2013; @kaspar_species-specific_nodate; @alexander_potential_2019; @finzi_carbon_2020
---

@@

```{r , eval=FALSE, echo=FALSE}
## Tallies for the following paragraph (Dec.21)
#sameV_sameW: LDWp, NMt, NIOt, NIOp, SCBIp, SCt, ZOFp
#sameV_diffW: BCNMp, BCNMt, CBp, CBt, HFt, HKKp, LDWt, NMp, SCBIt, SCp, ZOFt
#diffV_diffW: HFp, HKKt
```
The three metrics of growth ($RW$, $BAI$, and $\Delta AGB$)  exhibited similar strength of correlation and direction of response to climate variables across the range of potential time windows, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).
In 7 of 20 cases, both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_sameV_sameW`). 
In 11 cases, *climwin* identified the same climate variable but different (often overlapping) time windows (e.g., Fig. `r climwin_sameV_diffW`). 
Finally, in two cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at HKK, precipitation variable group at HF; Figs. `r clim_sens`, `r GLS_HKK`, `r GLS_HF`), *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_diffV_diffW`). 
Henceforth, unless otherwise noted, we focus on the climate sensitivities identified with $RW$ as the growth metric and for the full set of cores (*i.e.*, including those for which $DBH$ could not be reconstructed).

```{r , eval=FALSE, echo=FALSE}
## Tallies for the paragraph below (as of Dec. 22)
#BCNM: 1 species with both, 2 with just P
#HKK: 3 species with both, 1 with just T
#SCBI: 11 species with both, 3 with just P (FRNI, FAGR, PIST)
#LDW:  4 species with both,  2  with just P
#HF:  4 species with both
#ZOF:  3 species with both
#NIO:  1 species with both
#LT:  2 species with both
#CB:  7 species with both,  1 with just P (POTR)
#SC:  1 species with both
```
Within the context of the GLS model, which tested for first- and negative second- order linear effects of both a precipitation and a temperature variable, both a precipitation and a temperature variable were included in the top model for 80% of site-species combinations (n=37 of 46; Fig. `r clim_sens`). 
There were eight site-species combinations for which only a precipitation term was significant (2 at BCNM, 3 at SCBI, 2 at LDW, and 1 at CB), one for which only a temperature term was significant (CHTA at HKK), and none with no significant climate responses. 
Precipitation and temperature variables were rarely influential over the same time window (exception: LDW).
Below, we summarize the precipitation and temperature variables included in these models and their direction of response.

@@

###Windows selected: 
# BCNM: precip dry season (12mo), T dry season (2 mo) 
#HKK: PDF dry season + edges (9mo), T wet season (8 mo)
#SCBI: precip p June - cAug (15 mo), PET current gs (3 mo)
#LDW: precip c June (1 mo), PET current June (1mo)
#HF: PDF pMay (1mo), PET cMar (1mo)
#ZOF: pre pJune-pJul (2mo), TMX Jan-Mar (3mo)
#NIO: wet cMay (1mo), TMP pJune (1mo)
#LT: P pNov-cJul (9mo), TMX cMay-Jul (3mo)
#CB: P cApr-cJul(4mo), TMXpJul-cJul(13mo)
#SC: P pJun-pAug(3mo), TMXcJul (1mo)
```
Responses to precipitation amount ($PPT$) and frequency ($PDF$) were included in the best model for all but one species, and were predominantly positive (Fig. `r clim_sens`).
$PPT$ and $PDF$ were each identified as the top variable in four of the eight sites with both variables available, while $PPT$ was the only option at two (LT and CB).
Optimal time windows most commonly coincided at least partially with the months of most active growth in the current year (Fig. `r clim_sens`; Table `r site_details`):  wet seasons in the tropics (BCNM and HKK) or late spring/ early summer in the extratropics (n= 5 of 8).
In the tropics, the long time windows over which precipitation was influential (12 mo at BCNM, 9 mo at HKK) also included the majority (BCNM) or all (HKK) of the dry season.
At three extratropical sites (LDW, NIO, and CB), the most influential windows were limited to the current spring and early summer, whereas optimal windows at two sites extended back to the previous fall (LT) or summer (SCBI).
Finally, at three sites (HF, ZOF, and SC), precipitation of the previous growing season was the most influential variable. 
Responses to $PPT$ or $PDF$ were predominantly positive, with positive first-order linear terms for all but two species (*Tsuga canadensis* at HF and *Betula papyrifera* at NIO; Fig. `r clim_sens`).
Negative second-order terms were commonly included in the best model (*# of #* with positive first-order terms), generally resulting in a deceleration or decline at the highest levels of precipitation, but occasionally producing a unimodal (e.g., several species at SCBI) or predominantly negative response (e.g., BEAL at HF; Fig. `r clim_sens`). 
At the riparian site (NIO), for which streamflow was included as a candidate variable in the precipitation group, May $PDF$ was selected as the strongest driver but had a modest negative effect on growth, indicating that water availability was not a significant limitation for *Betula papyrifera* at this site.

A temperature variable was included in the best model for all but eight site-species combinations, with predominantly negative responses (*#*% of species; Fig. `r clim_sens`).
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, which were identified by *climwin* as the top temperature-related driver at five and three of the ten sites, respectively, noting that $PET$ was not available for two sites. 
$T_{mean}$ and $T_{min}$ were each identified as the top driver at only one site (NIO and BCNM, respectively), noting that the effects of $T_{min}$ at BCNM were only marginally significant (Fig. `r clim_sens`).
Optimal time windows most commonly occurred during the peak growing season of the current year (n= 5 of 10 sites), but there were cases where optimal windows occurred during the preceding dry season (BCNM), late winter/early spring (HF, ZOF), the previous growing season (NIO), or a 13-month window extending from past to current growing season (CB).
Within the tropics, there was minimal effect of temperature at BCNM and a negative effect of wet season $T_{max}$ for three of four species at HKK.
For temperate sites with optimal time windows covering the current and/or past growing season, responses were universally negative (i.e., negative first-order linear or unimodal, peaking at temperatures lower than the long-term mean). 
In contrast, there were positive effects of Jan-March $T_{max}$ for all three species at ZOF and of March $PET$ for *Tsuga Canadensis* at HF, the latter contrasting with a negative response of the other three species at HF (Fig. `r clim_sens`).
At the highest-latitude site (SC), which has undergone rapid warming and permafrost melt, *Picea mariana* responded positively (but with wide 95% CI on the slope) to temperature over the full analysis period (1903-2013); however, responses were predominately positive prior to 1970 and predominantly negative afterwards (Fig. **CREATE AN SI FIGURE!**).

@@

**All three growth metrics, $RW$, $BAI$, and $\Delta AGB$, varied with $DBH$ for most species at all sites (Fig. `r size`).** 
$RW$ varied significantly with $DBH$ for `r round(n_spst_DBH_sig/ n_spst_DBH_analyzed*100)`% of species-site combinations (n= `r n_spst_DBH_sig` of `r n_spst_DBH_analyzed`), and the majority of best models included a significant second-order linear $DBH$ term (n= `r n_spst_DBH_2nd_order`, `r n_spst_DBH_2nd_order_neg` of which were negative; Table **S#**). 
There was substantial variation in these trends, with patterns mixed across both forests and species within a single stand (Fig. `r size`).
On one end of the spectrum, some species exhibited maximum $RW$ at low $DBH$, followed by fairly rapid declines in $RW$ with increasing $DBH$. 
Species following this pattern either had low to at most intermediate shade tolerance (e.g., *Melia azedarach* at HKK, *Juglans nigra* at SCBI, *Acer rubrum* at HF, *Betula papyrifera* at NIO; *Populous tremuloides* at CB; Table `r species_table`) and/or grew in relatively open stands (e.g., both species at LT, *Picea mariana* at SC; Fig. `r size`).
At the other end of the spectrum, some species had low $RW$ at small $DBH$, increased to peak $RW$ at intermediate $DBH$, and subsequently declined.
These included, but were not limited to, shade-tolerant species at sites where they presumably established under closed-canopy conditions (e.g., *Trichilia tuberculata* and *Tetragastris panamensis* at BCNM; *Fagus* spp. at SCBI and Žofín, *Acer saccharum* at LDW, *Picea* spp. at Žofín and CB; Table `r species_table`).

@@

When a precipitation variable, a temperature variable, and $DBH$ were all included as candidate variables in the GLS models, typically all three were included in the top model, regardless of the growth metric used.
Climate responses were generally similar to those described above for models without a $DBH$ term, although some of the weaker climate responses were not consistently included in top models (e.g., $T_{min}$ responses at BCNM; Figs. `r clim_sens`, `r all_GLS`).
Responses to $DBH$ were as described above (Fig. `r size`).

In general, $DBH$ explained more variation in growth rates than did climate, but its relative importance varied across growth metrics and sites (Figs. `r all_GLS`).
The relative importance of $DBH$ tended to be least for $RW$, intermediate for $BAI$, and highest for $\Delta AGB$ (e.g., at SCBI; Fig. `r GLS_SCBI`).
However, there were exceptions, particularly when $RW$ decreased steeply with $DBH$ (e.g., LT; Fig. `r GLS_LT`).
The relative importance of climate was modest (but still significant) at sites including SCBI (Fig. `r GLS_SCBI`), HF (Fig. `r GLS_HF`), and SC (Fig. `r GLS_SC`), and stronger at sites including BCNM (Fig. `r GLS_BCNM`) and LT (Fig. `r GLS_LT`).

@@

Main bias would be stand dynamics. This is probably the case at secondary forests (HF, SCBI...). Even within mature forests, light-demanding species that establish in gaps (JACO at BCNM, MEAZ at HKK) would tend to experience an increasingly competitive environment.

### 2020-12-30 doc/manuscript/MEE_manuscript_SI.Rmd
BCI-->BCNM, site table updates
#100

GLS_BCI_legend = "Figure S9. Best GLS models for Barro Colorado Island (Panama)"
GLS_HKK_legend = "Figure S10. Best GLS models for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S11. Best GLS models for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S12. Best GLS models for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S13. Best GLS models for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S14. Best GLS models for Zofin Forest (Czech Republic)"
GLS_NE_legend = "Figure S15. Best GLS models for Niobrara/ Halsey (Nebraska, USA)"
GLS_LT_legend = "Figure S16. Best GLS models for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S17. Best GLS models for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S18. Best GLS models for Scotty Creek (Northwest Territory, Canada)"

@@

### Barro Colorado Island, Panama

@@

*For BCI, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCI starting in 1929 [@paton_barro_2019].*

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
S2 <- read.csv("tables_figures/species.csv", col.names = c("species code", "family","latin name", "sites sampled", "leaf type", "leaf phenology", "light requirements*", "bark allometry**"), stringsAsFactors = FALSE, check.names = FALSE)
kable(S2, format = "latex", booktabs = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
  row_spec(0, italic = F, bold = T) %>%
  column_spec(8, width = "5cm") %>%
  column_spec(3, italic = T) %>%
  column_spec(8, italic = T) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) 
```
*For extratropical species, light requirements are classified based on the database of @niinemets_tolerance_2006. For tropical species, categorization is based on @alfaro-sanchez_growth_2017 for BCI and @vlam_temperature_2014 for HKK.

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
breakrow1<-15
breakrow2<-29
breakrow3<-42
col_width<-"2.5cm"
table_pt1 <- read.csv("tables_figures/qualitative_results_comparison.csv", 
                     colClasses=c("NULL", NA, NA, NA, NA, NA, NA), 
                     col.names=c("NULL", "species", "previously observed", "observed here", "previously observed", "observed here", "reference"), 
                     skip=0, nrows=breakrow1,
                     stringsAsFactors = FALSE, check.names = FALSE)
kable(table_pt1, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
  add_header_above(c(" ", "Precipitation response" = 2, "Temperature response" = 2)) %>%
  column_spec(2:5, width = col_width) %>%
  column_spec(6, width = "2cm") %>%
  pack_rows("Barro Colorado Island, Panama", 1, 3, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  pack_rows("Huai Kha Khaeng, Thailand", 4, 7, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  pack_rows("Smithsonian Conservation Biology Institute, Virginia, USA", 8, 15, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  kable_styling(latex_options = c( "hold_position", "repeat header"), protect_latex = T, font_size = 7) 
# refs in this table need to go in the nocite field below
```

@@

\newpage
`r qualitative_results_comparison`, cont.
```{r eval = TRUE, echo=FALSE, warning=FALSE}
table_pt3 <- read.csv("tables_figures/qualitative_results_comparison.csv", 
                     colClasses=c("NULL", NA, NA, NA, NA, NA, NA), 
                     col.names=c("NULL", "species", "previously observed", "observed here", "previously observed", "observed here", "reference"), 
                     skip=breakrow2, nrows=breakrow3-breakrow2,
                     stringsAsFactors = FALSE, check.names = FALSE)
kable(table_pt3, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
  add_header_above(c(" ", "Precipitation response" = 2, "Temperature response" = 2)) %>%
  column_spec(2:5, width = col_width) %>%
  column_spec(6, width = "2cm") %>%
  pack_rows("Žofín Forest Dynamics Plot, Czech Republic", 1, 4, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  pack_rows("Niobrara and Hansley, Nebraska, USA", 5, 7, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  #pack_rows("Scotty Creek, NW Territories, Canada", 13, 13, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  kable_styling(latex_options = c( "hold_position", "repeat header"), protect_latex = T, font_size = 7) 
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @alfaro-sanchez_growth_2017; @vlam_temperature_2014; @helcoski_growing_2019; @maxwell_declining_2016; @aus_de_ar_tree_2018; @bumann_assessing_2019; @tumajer_increasing_2017; @sniderhan_growth_2016; @kaspar_species-specific_nodate
---
*Indicates results from a regional study including but not limited to cores from the focal site.

@@

\newpage
## `r GLS_BCI_legend`

![`r GLS_BCI_legend` `r GLS_generic_legend`](tables_figures/SI_figures/composite_plots/BCI.png){ width=75% }

>>>>

GLS_BCNM_legend = "Figure S9. Best GLS models for Barro Colorado Nature Monument (Panama)"
GLS_HKK_legend = "Figure S10. Best GLS models for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S11. Best GLS models for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S12. Best GLS models for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S13. Best GLS models for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S14. Best GLS models for Zofin Forest (Czech Republic)"
GLS_NE_legend = "Figure S15. Best GLS models for Niobrara/ Halsey (Nebraska, USA)"
GLS_LT_legend = "Figure S16. Best GLS models for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S17. Best GLS models for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S18. Best GLS models for Scotty Creek (Northwest Territory, Canada)"

@@

### Barro Colorado Nature Monument, Panama

@@

*For BCNM, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCNM starting in 1929 [@paton_barro_2019].*

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
S2 <- read.csv("tables_figures/species.csv", col.names = c("species code", "family","latin name", "sites sampled", "leaf type", "leaf phenology", "light requirements*", "bark allometry**"), stringsAsFactors = FALSE, check.names = FALSE)
kable(S2, format = "latex", booktabs = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
  row_spec(0, italic = F, bold = T) %>%
  column_spec(8, width = "5cm") %>%
  column_spec(3, italic = T) %>%
  column_spec(8, italic = T) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) 
```
*For extratropical species, light requirements are classified based on the database of @niinemets_tolerance_2006. For tropical species, categorization is based on @alfaro-sanchez_growth_2017 for BCNM and @vlam_temperature_2014 for HKK.

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
breakrow1<-15
breakrow2<-29
breakrow3<-42
col_width<-"2.5cm"
table_pt1 <- read.csv("tables_figures/qualitative_results_comparison.csv", 
                     colClasses=c("NULL", NA, NA, NA, NA, NA, NA), 
                     col.names=c("NULL", "species", "previously observed", "observed here", "previously observed", "observed here", "reference"), 
                     skip=0, nrows=breakrow1,
                     stringsAsFactors = FALSE, check.names = FALSE)
kable(table_pt1, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
  add_header_above(c(" ", "Precipitation response" = 2, "Temperature response" = 2)) %>%
  column_spec(2:5, width = col_width) %>%
  column_spec(6, width = "2cm") %>%
  pack_rows("Barro Colorado Nature Monument, Panama", 1, 3, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  pack_rows("Huai Kha Khaeng, Thailand", 4, 7, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  pack_rows("Smithsonian Conservation Biology Institute, Virginia, USA", 8, 15, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  kable_styling(latex_options = c( "hold_position", "repeat header"), protect_latex = T, font_size = 7) 
# refs in this table need to go in the nocite field below
```

@@

\newpage
`r qualitative_results_comparison`, cont.
```{r eval = TRUE, echo=FALSE, warning=FALSE}
table_pt3 <- read.csv("tables_figures/qualitative_results_comparison.csv", 
                     colClasses=c("NULL", NA, NA, NA, NA, NA, NA), 
                     col.names=c("NULL", "species", "previously observed", "observed here", "previously observed", "observed here", "reference"), 
                     skip=breakrow2, nrows=breakrow3-breakrow2,
                     stringsAsFactors = FALSE, check.names = FALSE)
kable(table_pt3, format = "latex", booktabs = TRUE, longtable = TRUE,  escape = F, linesep = c('\\addlinespace'))%>%
  add_header_above(c(" ", "Precipitation response" = 2, "Temperature response" = 2)) %>%
  column_spec(2:5, width = col_width) %>%
  column_spec(6, width = "2cm") %>%
  pack_rows("Žofín Forest Dynamics Plot, Czech Republic", 1, 4, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  pack_rows("Niobrara and Hansley, Nebraska, USA", 5, 7, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  #pack_rows("Scotty Creek, NW Territories, Canada", 13, 13, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
  kable_styling(latex_options = c( "hold_position", "repeat header"), protect_latex = T, font_size = 7) 
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @alfaro-sanchez_growth_2017; @vlam_temperature_2014; @helcoski_growing_2019; @maxwell_declining_2016; @bumann_assessing_2019; @tumajer_increasing_2017; @sniderhan_growth_2016; @kaspar_species-specific_nodate
---
*Indicates results from a regional study including but not limited to cores from the focal site.

@@

\newpage
## `r GLS_BCNM_legend`

![`r GLS_BCNM_legend` `r GLS_generic_legend`](tables_figures/SI_figures/composite_plots/BCI.png){ width=75% }
